% \documentclass[ba,preprint]{imsart}% use this for supplement article
\documentclass[ba]{imsart}
\input{definitions}

%% Packages
\RequirePackage{amsthm,amsmath,amsfonts,amssymb}
\RequirePackage[numbers]{natbib}
%\RequirePackage[authoryear]{natbib}%% uncomment this for author-year citations
\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue,backref=page,backref=page]{hyperref}
\RequirePackage{graphicx}

\pubyear{2024}
\arxiv{2007.12032}
\volume{TBA}
\issue{TBA}
\firstpage{1}
\lastpage{1}

\startlocaldefs
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Uncomment next line to change            %%
%% the type of equation numbering           %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\numberwithin{equation}{section}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% For Axiom, Claim, Corollary, Hypothesis, %%
%% Lemma, Theorem, Proposition              %%
%% use \theoremstyle{plain}                 %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\theoremstyle{plain}
\newtheorem{axiom}{Axiom}
\newtheorem{claim}[axiom]{Claim}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% For Assumption, Definition, Example,     %%
%% Notation, Property, Remark, Fact         %%
%% use \theoremstyle{definition}            %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem*{example}{Example}
\newtheorem*{fact}{Fact}
\newtheorem*{ctheorem}{Theorem}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% For Case use \theoremstyle{remark}       %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\theoremstyle{remark}
\newtheorem{case}{Case}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Please put your definitions here:        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\endlocaldefs


\begin{document}

\begin{frontmatter}
  \title{Measurement Clusterization in Bayesian D-optimal Designs in Infinite Dimensions}
  \runtitle{Measurement Clusterization in D-optimal Designs}
  

\begin{aug}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Only one address is permitted per author. %%
%% Only division, organization and e-mail is %%
%% included in the address.                  %%
%% Additional information can be included in %%
%% the Acknowledgments section if necessary. %%
%% ORCID can be inserted by command:         %%
%% \orcid{0000-0000-0000-0000}               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author[A]{\fnms{Yair}~\snm{Daon}\ead[label=e1]{yair.daon@gmail.com}\orcid{0000-0002-7491-2008}}
%% \author[A]{\fnms{Second}~\snm{Author}\ead[label=e2]{second@somewhere.com}
%% \and
%% \author[B]{\fnms{Third}~\snm{Author}\ead[label=e3]{third@somewhere.com}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Addresses                                %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\address[A]{Azrieli Faculty of Medicine, Bar-Ilan University 
\printead[presep={,\ }]{e1}}

%% \address[B]{Department,
%% University or Company Name\printead[presep={,\ }]{e2,e3}}
%% \runauthor{F. Author et al.}
\end{aug}

\begin{abstract}
  Estimation of parameters in physical processes often demands costly
  measurements, prompting the pursuit of an optimal measurement
  strategy. Finding such strategy is termed the problem of optimal
  experimental design, abbreviated as optimal design. Remarkably,
  optimal designs can yield tightly clustered measurement locations,
  leading researchers to fundamentally revise the design problem just
  to circumvent this issue. Some authors introduce error correlation
  among error terms that are initially independent, while others
  restrict measurement locations to a finite set of locations. While
  both approaches may prevent clusterization, they also fundamentally
  alter the optimal design problem.

  In this study, we consider Bayesian D-optimal designs, i.e.~designs
  that maximize the expected Kullback-Leibler divergence between
  posterior and prior. We propose an analytically tractable model for
  D-optimal designs over Hilbert spaces. In this framework, we make
  several key contributions: (a) We establish that
  measurement clusterization is a generic trait of D-optimal designs
  for linear inverse problems with independent Gaussian measurement
  errors and a Gaussian prior. (b) We prove that introducing
  correlations among measurement error terms mitigates
  clusterization.(c) We characterize D-optimal designs as
  reducing uncertainty across a subset of prior covariance
  eigenvectors. (d) We leverage this characterization to
  argue that measurement clusterization arises as a consequence of the
  pigeonhole principle: when more measurements are taken than there
  are locations where the select eigenvectors are large and others are
  small --- clusterization occurs. Finally, we use our analysis to
  argue against the use of Gaussian priors with linearized physical
  models when seeking a D-optimal design.
\end{abstract}

%% ** Keywords **
\begin{keyword}[class=MSC]
\kwd[Primary ]{62F15}
\kwd{35R30}
\kwd[; secondary ]{28C20}
\end{keyword}

\begin{keyword}
\kwd{D-optimal design}
\kwd{inverse problem}
\kwd{clusterization}
\end{keyword}

\end{frontmatter}

\input{intro}
\input{prelims}
\input{constrained}
\input{nonvanishing}
\input{vanishing}
\input{numerix}
%\input{discussion}


\begin{acks}[Acknowledgments]
  This study is a result of research I started during my PhD studies
  under the instruction of Prof.~Georg Stadler at the Courant
  Institute of Mathematical Sciences. I would like to thank him for
  his great mentorship, attention to details and kindness. I would
  also like to thank Christian Remling, who helped me find a proof for
  Lemma \ref{lemma:free} in
  \href{https://mathoverflow.net/questions/280168/redistribute-diagonal-entries-of-a-matrix/280203#280203c}{Mathoverflow}.
  Last, but certainly not least, I would like to thank the three
  referees, associate editor and editor in chief for providing
  detailed and insightful reviewes that have made this manuscript a
  whole lot better.
  %Having worked on this manuscript alone for several years, they have
  %almost become my coautoh
\end{acks}

\begin{funding} 
  This research was supported in part by an appointment with the
  National Science Foundation (NSF) Mathematical Sciences Graduate
  Internship (MSGI) Program sponsored by the NSF Division of
  Mathematical Sciences. This program is administered by the Oak Ridge
  Institute for Science and Education (ORISE) through an interagency
  agreement between the U.S. Department of Energy (DOE) and NSF. ORISE
  is managed for DOE by ORAU. All opinions expressed in this paper are
  the author's and do not necessarily reflect the policies and views
  of NSF, ORAU/ORISE, or DOE.
  % This work was also supported by The Raymond and Beverly Sackler
  % Post-Doctoral Scholarship.
\end{funding}

\bibliographystyle{ba}
\bibliography{~/lib.bib}

\end{document}

